---
title: "Random Forest Regressions"
format: html
editor: visual
toc: true
toc-depth: 4
df-print: paged
code-overflow: wrap
---

This script is for outcome of interest: total reported cases per week

# Load Packages

```{r,echo=FALSE}
knitr::opts_chunk$set(message=F)
```

```{r,message=FALSE}
library(tidyverse)
library(here)
library(tidymodels)
library(ggpubr)
library(tsibble)
library(ingredients)
library(RColorBrewer)
```

# Load Data

```{r}
# n = 6 (original data)
data_n6 <- readRDS(here("data/processed_data/wbe_covid_n6_week.rds")) %>% 
  drop_na() %>% 
  mutate_at(vars(14:19),log10) %>% 
  mutate(A_N1_POS = n_pos_A_N1/n_A_N1,
         A_N2_POS = n_pos_A_N2/n_A_N2,
         B_N1_POS = n_pos_B_N1/n_B_N1,
         B_N2_POS = n_pos_B_N2/n_B_N2,
         C_N1_POS = n_pos_C_N1/n_C_N1,
         C_N2_POS = n_pos_C_N2/n_C_N2,
         log10_cases = log10(cases.reported))

# n = 5
data_n5 <- readRDS(here("data/processed_data/wbe_covid_n5_week.rds")) %>% 
  drop_na() %>% 
  mutate_at(vars(14:19),log10) %>% 
  mutate(A_N1_POS = n_pos_A_N1/n_A_N1,
         A_N2_POS = n_pos_A_N2/n_A_N2,
         B_N1_POS = n_pos_B_N1/n_B_N1,
         B_N2_POS = n_pos_B_N2/n_B_N2,
         C_N1_POS = n_pos_C_N1/n_C_N1,
         C_N2_POS = n_pos_C_N2/n_C_N2,
         log10_cases = log10(cases.reported))

# n = 4
data_n4 <- readRDS(here("data/processed_data/wbe_covid_n4_week.rds")) %>% 
  drop_na() %>% 
  mutate_at(vars(14:19),log10) %>% 
  mutate(A_N1_POS = n_pos_A_N1/n_A_N1,
         A_N2_POS = n_pos_A_N2/n_A_N2,
         B_N1_POS = n_pos_B_N1/n_B_N1,
         B_N2_POS = n_pos_B_N2/n_B_N2,
         C_N1_POS = n_pos_C_N1/n_C_N1,
         C_N2_POS = n_pos_C_N2/n_C_N2,
         log10_cases = log10(cases.reported))

# n = 3
data_n3 <- readRDS(here("data/processed_data/wbe_covid_n3_week.rds")) %>% 
  drop_na() %>% 
  mutate_at(vars(14:19),log10) %>% 
  mutate(A_N1_POS = n_pos_A_N1/n_A_N1,
         A_N2_POS = n_pos_A_N2/n_A_N2,
         B_N1_POS = n_pos_B_N1/n_B_N1,
         B_N2_POS = n_pos_B_N2/n_B_N2,
         C_N1_POS = n_pos_C_N1/n_C_N1,
         C_N2_POS = n_pos_C_N2/n_C_N2,
         log10_cases = log10(cases.reported))

# n = 2
data_n2 <- readRDS(here("data/processed_data/wbe_covid_n2_week.rds")) %>% 
  drop_na() %>% 
  mutate_at(vars(14:19),log10) %>% 
  mutate(A_N1_POS = n_pos_A_N1/n_A_N1,
         A_N2_POS = n_pos_A_N2/n_A_N2,
         B_N1_POS = n_pos_B_N1/n_B_N1,
         B_N2_POS = n_pos_B_N2/n_B_N2,
         C_N1_POS = n_pos_C_N1/n_C_N1,
         C_N2_POS = n_pos_C_N2/n_C_N2,
         log10_cases = log10(cases.reported))

# n = 1
data_n1 <- readRDS(here("data/processed_data/wbe_covid_n1_week.rds")) %>% 
  drop_na() %>% 
  mutate_at(vars(14:19),log10) %>% 
  mutate(A_N1_POS = n_pos_A_N1/n_A_N1,
         A_N2_POS = n_pos_A_N2/n_A_N2,
         B_N1_POS = n_pos_B_N1/n_B_N1,
         B_N2_POS = n_pos_B_N2/n_B_N2,
         C_N1_POS = n_pos_C_N1/n_C_N1,
         C_N2_POS = n_pos_C_N2/n_C_N2,
         log10_cases = log10(cases.reported))
```

# Split data

```{r}
data_n6_train <- data_n6 %>% head(n = 79)
data_n6_test <- data_n6 %>% tail(n = 52)

data_n5_train <- data_n5 %>% head(n = 79)
data_n5_test <- data_n5 %>% tail(n = 52)

data_n4_train <- data_n4 %>% head(n = 79)
data_n4_test <- data_n4 %>% tail(n = 52)

data_n3_train <- data_n3 %>% head(n = 79)
data_n3_test <- data_n3 %>% tail(n = 52)

data_n2_train <- data_n2 %>% head(n = 79)
data_n2_test <- data_n2 %>% tail(n = 52)

data_n1_train <- data_n1 %>% head(n = 79)
data_n1_test <- data_n1 %>% tail(n = 52)
```

# Specify models & workflows

```{r}
#general model for all iterations
rf_model <- rand_forest() %>% 
  set_engine("ranger", importance = "permutation") %>% 
  set_mode("regression") 

#workflow for assay pos models
ap_workflow <- workflow() %>%
  add_model(rf_model) %>% 
  add_formula(log10_cases ~ 
              A_N1_POS+A_N2_POS+B_N1_POS+B_N2_POS+C_N1_POS+C_N2_POS)

#workflow for viral load models
vl_workflow <- workflow() %>% 
  add_model(rf_model) %>% 
  add_formula(log10_cases ~ A_N1+A_N2+B_N1+B_N2+C_N1+C_N2)

#workflow for all predictors model
all_workflow <- workflow() %>% 
  add_model(rf_model) %>% 
  add_formula(log10_cases ~ 
              A_N1_POS+A_N2_POS+B_N1_POS+B_N2_POS+C_N1_POS+C_N2_POS+
              A_N1+A_N2+B_N1+B_N2+C_N1+C_N2)

#workflow for assay pos plant A
select_ap_workflow <- workflow() %>% 
  add_model(rf_model) %>% 
  add_formula(log10_cases ~ A_N1_POS+A_N2_POS)

#workflow for assay pos/viral load plant A
select_workflow <- workflow() %>% 
  add_model(rf_model) %>% 
  add_formula(log10_cases ~ A_N1_POS+A_N2_POS+A_N1+A_N2)
```

# Fit Detection Frequency

## n = 6

```{r}
#fit model to data
set.seed(13)
fit_ap_n6 <- ap_workflow %>% 
  fit(data = data_n6_train)

#create folds for cross validation
set.seed(13)
folds_ap_n6 <- vfold_cv(data_n6_train, 
                        v = 2, repeats = 20, strata = log10_cases)

#run cross validation
set.seed(13)
cv_ap_n6 <- fit_resamples(ap_workflow, resamples = folds_ap_n6)
cv_ap_n6_metrics <- collect_metrics(cv_ap_n6,summarize = F)
ap_n6_metrics_sum <- collect_metrics(cv_ap_n6,summarize = T)

#check cross validation metrics
ap_n6_metrics_sum 
```

```{r}
#fit model to training data
set.seed(13)
fit_ap_n6 <- ap_workflow %>% 
  fit(data = data_n6_train)

# Extract the fitted ranger model
rf_fitted_model_ap_n6 <- extract_fit_parsnip(fit_ap_n6)$fit

# Assess variable importance using vip package
var_importance_ap_n6 <- vip::vi(rf_fitted_model_ap_n6)

# Print variable importance
print(var_importance_ap_n6)

# Optional: Plot variable importance
vip::vip(rf_fitted_model_ap_n6)

ggsave(here("figures/rand_forest_importance/importance_df_predictors.png"))
```

## n = 5

```{r}
#fit model to data
set.seed(13)
fit_ap_n5 <- ap_workflow %>% 
  fit(data = data_n5_train)

#create folds for cross validation
set.seed(13)
folds_ap_n5 <- vfold_cv(data_n5_train, 
                        v = 2, repeats = 20, strata = log10_cases)

#run cross validation
set.seed(13)
cv_ap_n5 <- fit_resamples(ap_workflow, resamples = folds_ap_n5)
cv_ap_n5_metrics <- collect_metrics(cv_ap_n5,summarize = F)
ap_n5_metrics_sum <- collect_metrics(cv_ap_n5,summarize = T)

#check cross validation metrics
ap_n5_metrics_sum 
```

## n = 4

```{r}
#fit model to data
set.seed(13)
fit_ap_n4 <- ap_workflow %>% 
  fit(data = data_n4_train)

#create folds for cross validation
set.seed(13)
folds_ap_n4 <- vfold_cv(data_n4_train, 
                        v = 2, repeats = 20, strata = log10_cases)

#run cross validation
set.seed(13)
cv_ap_n4 <- fit_resamples(ap_workflow, resamples = folds_ap_n4)
cv_ap_n4_metrics <- collect_metrics(cv_ap_n4,summarize = F)
ap_n4_metrics_sum <- collect_metrics(cv_ap_n4,summarize = T)

#check cross validation metrics
ap_n4_metrics_sum
```

## n = 3

```{r}
#fit model to data
set.seed(13)
fit_ap_n3 <- ap_workflow %>% 
  fit(data = data_n3_train)

#create folds for cross validation
set.seed(13)
folds_ap_n3 <- vfold_cv(data_n3_train, 
                        v = 2, repeats = 20, strata = log10_cases)

#run cross validation
set.seed(13)
cv_ap_n3 <- fit_resamples(ap_workflow, resamples = folds_ap_n3)
cv_ap_n3_metrics <- collect_metrics(cv_ap_n3,summarize = F)
ap_n3_metrics_sum <- collect_metrics(cv_ap_n3,summarize = T)

#check cross validation metrics
ap_n3_metrics_sum 
```

## n = 2

```{r}
#fit model to data
set.seed(13)
fit_ap_n2 <- ap_workflow %>% 
  fit(data = data_n2_train)

#create folds for cross validation
set.seed(13)
folds_ap_n2 <- vfold_cv(data_n2_train, v = 2, repeats = 20, 
                        strata = log10_cases)

#run cross validation
set.seed(13)
cv_ap_n2 <- fit_resamples(ap_workflow, resamples = folds_ap_n2)
cv_ap_n2_metrics <- collect_metrics(cv_ap_n2,summarize = F)
ap_n2_metrics_sum <- collect_metrics(cv_ap_n2,summarize = T)

#check cross validation metrics
ap_n2_metrics_sum 
```

## n =1

```{r}
#fit model to data
set.seed(13)
fit_ap_n1 <- ap_workflow %>% 
  fit(data = data_n1_train)

#create folds for cross validation
set.seed(13)
folds_ap_n1 <- vfold_cv(data_n1_train, v = 2, repeats = 20, 
                        strata = log10_cases)

#run cross validation
set.seed(13)
cv_ap_n1 <- fit_resamples(ap_workflow, resamples = folds_ap_n1)
cv_ap_n1_metrics <- collect_metrics(cv_ap_n1,summarize = F)
ap_n1_metrics_sum <- collect_metrics(cv_ap_n1,summarize = T)

#check cross validation metrics
ap_n1_metrics_sum 
```

# Fit Viral Load

## n = 6

```{r}
#fit model to data
set.seed(13)
fit_vl_n6 <- vl_workflow %>% 
  fit(data = data_n6_train)

#create folds for cross validation
set.seed(13)
folds_vl_n6 <- vfold_cv(data_n6_train, v = 2, repeats = 20, 
                        strata = log10_cases)

#run cross validation
set.seed(13)
cv_vl_n6 <- fit_resamples(vl_workflow, resamples = folds_vl_n6)
cv_vl_n6_metrics <- collect_metrics(cv_vl_n6,summarize = F)

#check cross validation metrics
cv_vl_n6_metrics 
```

```{r}
#fit model to training data
set.seed(13)
fit_vl_n6 <- vl_workflow %>% 
  fit(data = data_n6_train)

# Extract the fitted ranger model
rf_fitted_model_vl_n6 <- extract_fit_parsnip(fit_vl_n6)$fit

# Assess variable importance using vip package
var_importance_vl_n6 <- vip::vi(rf_fitted_model_vl_n6)

# Print variable importance
print(var_importance_vl_n6)

# Optional: Plot variable importance
vip::vip(rf_fitted_model_vl_n6)

ggsave(here("figures/rand_forest_importance/importance_vl_predictors.png"))
```

## n = 5

```{r}
#fit model to data
set.seed(13)
fit_vl_n5 <- vl_workflow %>% 
  fit(data = data_n5_train)

#create folds for cross validation
set.seed(13)
folds_vl_n5 <- vfold_cv(data_n5_train, v = 2, repeats = 20, 
                        strata = log10_cases)

#run cross validation
set.seed(13)
cv_vl_n5 <- fit_resamples(vl_workflow, resamples = folds_vl_n5)
cv_vl_n5_metrics <- collect_metrics(cv_vl_n5,summarize = F)

#check cross validation metrics
cv_vl_n5_metrics 
```

## n = 4

```{r}
#fit model to data
set.seed(13)
fit_vl_n4 <- vl_workflow %>% 
  fit(data = data_n4_train)

#create folds for cross validation
set.seed(13)
folds_vl_n4 <- vfold_cv(data_n4_train, v = 2, repeats = 20, 
                        strata = log10_cases)

#run cross validation
set.seed(13)
cv_vl_n4 <- fit_resamples(vl_workflow, resamples = folds_vl_n4)
cv_vl_n4_metrics <- collect_metrics(cv_vl_n4,summarize = F)

#check cross validation metrics
cv_vl_n4_metrics 
```

## n = 3

```{r}
#fit model to data
set.seed(13)
fit_vl_n3 <- vl_workflow %>% 
  fit(data = data_n3_train)

#create folds for cross validation
set.seed(13)
folds_vl_n3 <- vfold_cv(data_n3_train, v = 2, repeats = 20, 
                        strata = log10_cases)

#run cross validation
set.seed(13)
cv_vl_n3 <- fit_resamples(vl_workflow, resamples = folds_vl_n3)
cv_vl_n3_metrics <- collect_metrics(cv_vl_n3,summarize = F)

#check cross validation metrics
cv_vl_n3_metrics 
```

## n = 2

```{r}
#fit model to data
set.seed(13)
fit_vl_n2 <- vl_workflow %>% 
  fit(data = data_n2_train)

#create folds for cross validation
set.seed(13)
folds_vl_n2 <- vfold_cv(data_n2_train, v = 2, repeats = 20, 
                        strata = log10_cases)

#run cross validation
set.seed(13)
cv_vl_n2 <- fit_resamples(vl_workflow, resamples = folds_vl_n2)
cv_vl_n2_metrics <- collect_metrics(cv_vl_n2,summarize = F)

#check cross validation metrics
cv_vl_n2_metrics 
```

## n = 1

```{r}
#fit model to data
set.seed(13)
fit_vl_n1 <- vl_workflow %>% 
  fit(data = data_n1_train)

#create folds for cross validation
set.seed(13)
folds_vl_n1 <- vfold_cv(data_n1_train, v = 2, repeats = 20, 
                        strata = log10_cases)

#run cross validation
set.seed(13)
cv_vl_n1 <- fit_resamples(vl_workflow, resamples = folds_vl_n1)
cv_vl_n1_metrics <- collect_metrics(cv_vl_n1,summarize = F)

#check cross validation metrics
cv_vl_n1_metrics 
```

# Fit all predictors

## n = 6

```{r}
#fit model to training data
set.seed(13)
fit_all_n6 <- all_workflow %>% 
  fit(data = data_n6_train)

# Extract the fitted ranger model
rf_fitted_model_all_n6 <- extract_fit_parsnip(fit_all_n6)$fit

# Assess variable importance using vip package
var_importance_all_n6 <- vip::vi(rf_fitted_model_all_n6)

# Print variable importance
print(var_importance_all_n6)

# Optional: Plot variable importance
vip::vip(rf_fitted_model_all_n6)

ggsave(here("figures/rand_forest_importance/importance_all_predictors.png"))
```

```{r}
#fit model to data
set.seed(13)
fit_all_n6 <- all_workflow %>% 
  fit(data = data_n6_train)

#create folds for cross validation
set.seed(13)
folds_all_n6 <- vfold_cv(data_n6_train, v = 2, repeats = 20, 
                        strata = log10_cases)

#run cross validation
set.seed(13)
cv_all_n6 <- fit_resamples(all_workflow, resamples = folds_all_n6)
cv_all_n6_metrics <- collect_metrics(cv_all_n6,summarize = F)

#check cross validation metrics
cv_all_n6_metrics 
```

## n = 5

```{r}
#fit model to data
set.seed(13)
fit_all_n5 <- all_workflow %>% 
  fit(data = data_n5_train)

#create folds for cross validation
set.seed(13)
folds_all_n5 <- vfold_cv(data_n5_train, v = 2, repeats = 20, 
                        strata = log10_cases)

#run cross validation
set.seed(13)
cv_all_n5 <- fit_resamples(all_workflow, resamples = folds_all_n5)
cv_all_n5_metrics <- collect_metrics(cv_all_n5,summarize = F)

#check cross validation metrics
cv_all_n5_metrics 
```

## n = 4

```{r}
#fit model to data
set.seed(13)
fit_all_n4 <- all_workflow %>% 
  fit(data = data_n4_train)

#create folds for cross validation
set.seed(13)
folds_all_n4 <- vfold_cv(data_n4_train, v = 2, repeats = 20, 
                        strata = log10_cases)

#run cross validation
set.seed(13)
cv_all_n4 <- fit_resamples(all_workflow, resamples = folds_all_n4)
cv_all_n4_metrics <- collect_metrics(cv_all_n4,summarize = F)

#check cross validation metrics
cv_all_n4_metrics 
```

## n = 3

```{r}
#fit model to data
set.seed(13)
fit_all_n3 <- all_workflow %>% 
  fit(data = data_n3_train)

#create folds for cross validation
set.seed(13)
folds_all_n3 <- vfold_cv(data_n3_train, v = 2, repeats = 20, 
                        strata = log10_cases)

#run cross validation
set.seed(13)
cv_all_n3 <- fit_resamples(all_workflow, resamples = folds_all_n3)
cv_all_n3_metrics <- collect_metrics(cv_all_n3,summarize = F)

#check cross validation metrics
cv_all_n3_metrics 
```

## n = 2

```{r}
#fit model to data
set.seed(13)
fit_all_n2 <- all_workflow %>% 
  fit(data = data_n2_train)

#create folds for cross validation
set.seed(13)
folds_all_n2 <- vfold_cv(data_n2_train, v = 2, repeats = 20, 
                        strata = log10_cases)

#run cross validation
set.seed(13)
cv_all_n2 <- fit_resamples(all_workflow, resamples = folds_all_n2)
cv_all_n2_metrics <- collect_metrics(cv_all_n2,summarize = F)

#check cross validation metrics
cv_all_n2_metrics 
```

## n = 1

```{r}
#fit model to data
set.seed(13)
fit_all_n1 <- all_workflow %>% 
  fit(data = data_n1_train)

#create folds for cross validation
set.seed(13)
folds_all_n1 <- vfold_cv(data_n1_train, v = 2, repeats = 20, 
                        strata = log10_cases)

#run cross validation
set.seed(13)
cv_all_n1 <- fit_resamples(all_workflow, resamples = folds_all_n1)
cv_all_n1_metrics <- collect_metrics(cv_all_n1,summarize = F)

#check cross validation metrics
cv_all_n1_metrics 
```

# Fit Selected Variables

## Detection Freq Plant A, n = 6

```{r}
#fit model to data
set.seed(13)
fit_select_ap_n6 <- select_ap_workflow %>% 
  fit(data = data_n6_train)

#create folds for cross validation
set.seed(13)
folds_select_ap_n6 <- vfold_cv(data_n6_train, v = 2, repeats = 20, 
                        strata = log10_cases)

#run cross validation
set.seed(13)
cv_select_ap_n6 <- fit_resamples(select_ap_workflow, 
                                 resamples = folds_select_ap_n6)
cv_select_ap_n6_metrics <- collect_metrics(cv_select_ap_n6,summarize = F)

#check cross validation metrics
cv_select_ap_n6_metrics 
```

## DF & VL Plant A, n = 6

```{r}
#fit model to data
set.seed(13)
fit_select_n6 <- select_workflow %>% 
  fit(data = data_n6_train)

#create folds for cross validation
set.seed(13)
folds_select_n6 <- vfold_cv(data_n6_train, v = 2, repeats = 20, 
                        strata = log10_cases)

#run cross validation
set.seed(13)
cv_select_n6 <- fit_resamples(select_workflow, 
                                 resamples = folds_select_n6)
cv_select_n6_metrics <- collect_metrics(cv_select_n6,summarize = F)

#check cross validation metrics
cv_select_n6_metrics 
```

# Augment Assay Pos Fits

## n = 6

### Training data

```{r}
set.seed(13)
aug_ap_n6 <- augment(fit_ap_n6, data_n6_train)

aug_ap_n6 %>% ggplot(aes(log10_cases, .pred)) +
  geom_point() +
  stat_smooth(method = "lm")

ggsave(here("figures/randforest_df_n6_obsVSpredict.png"))

df_time <- aug_ap_n6 %>% ggplot(aes(x=week)) +
  geom_smooth(aes(y=.pred), method = "gam", color="red") +
  geom_smooth(aes(y=log10_cases), method = "gam", color="blue") +
  ggthemes::theme_clean() +
  xlab("Week") +
  ylab("Test positivity rate")

df_time

ggsave(here("figures/randforest_df_n6_timeseries.png"))
```

### Testing data

```{r}
set.seed(13)
aug_ap_n6_test <- augment(fit_ap_n6, data_n6_test)

aug_ap_n6_test %>% ggplot(aes(log10_cases, .pred)) +
  geom_point() +
  stat_smooth(method = "lm") 

ggsave(here("figures/randforest_df_n6_obsVSpredict_2022.png"))

aug_ap_n6_test %>% ggplot(aes(x=week)) +
  geom_smooth(aes(y=.pred), method = "gam", color="red") +
  geom_smooth(aes(y=log10_cases), method = "gam", color="blue") +
  ggthemes::theme_clean() +
  ylab("Log10 reported cases per week")

ggsave(here("figures/randforest_df_n6_timeseries_2022.png"))
```

## n = 1

### Training data

```{r}
set.seed(13)
aug_ap_n1 <- augment(fit_ap_n1, data_n1_train)

aug_ap_n1 %>% ggplot(aes(log10_cases, .pred)) +
  geom_point() +
  stat_smooth(method = "lm")

df_n1_train <- aug_ap_n1 %>% ggplot(aes(x=week)) +
  geom_line(aes(y=log10_cases), color = "blue") +
  geom_line(aes(y=.pred), color = "red") +
  geom_smooth(aes(y=.pred), method = "gam", color="red") +
  geom_smooth(aes(y=log10_cases), method = "gam", color="blue") +
  ggthemes::theme_clean() +
  ylab("Clinical test positivity rate")

df_n1_train
```

### Testing data

```{r}
set.seed(13)
aug_ap_n1_test <- augment(fit_ap_n1, data_n1_test)

aug_ap_n1_test %>% ggplot(aes(log10_cases, .pred)) +
  geom_point() +
  stat_smooth(method = "lm") 
  

df_n1_test <- aug_ap_n1_test %>% ggplot(aes(x=week)) +
  geom_line(aes(y=log10_cases), color = "blue") +
  geom_line(aes(y=.pred), color = "red") +
  geom_smooth(aes(y=.pred), method = "gam", color="red") +
  geom_smooth(aes(y=log10_cases), method = "gam", color="blue") +
  ggthemes::theme_clean() +
  ylab("Clinical test positivity rate")

df_n1_test
```

```{r}
df_n1_time <- cowplot::plot_grid(df_n1_train,df_n1_test,nrow = 2)

saveRDS(df_n1_time, here("figures/project_8040/df_n1_predictions_timeseries.rds"))
```

# Augment Viral Load Fits

## n = 6

### Training data

```{r}
set.seed(13)
aug_vl_n6 <- augment(fit_vl_n6, data_n6_train)

aug_vl_n6 %>% ggplot(aes(log10_cases, .pred)) +
  geom_point() +
  stat_smooth(method = "lm")

ggsave(here("figures/randforest_vl_n6_obsVSpredict.png"))

vl_time <- aug_vl_n6 %>% ggplot(aes(x=week)) +
  geom_line(aes(y=log10_cases), color = "blue") +
  geom_line(aes(y=.pred), color = "red") +
  ggthemes::theme_clean() +
  xlab("Week") +
  ylab("Test positivity rate")

vl_time

ggsave(here("figures/randforest_vl_n6_timeseries.png"))
```

### Testing data

```{r}
set.seed(13)
aug_vl_n6_test <- augment(fit_vl_n6, data_n6_test)

aug_vl_n6_test %>% ggplot(aes(log10_cases, .pred)) +
  geom_point() +
  stat_smooth(method = "lm") 

ggsave(here("figures/randforest_vl_n6_obsVSpredict_2022.png"))

aug_vl_n6_test %>% ggplot(aes(x=week)) +
  geom_line(aes(y=log10_cases), color = "blue") +
  geom_line(aes(y=.pred), color = "red")

ggsave(here("figures/randforest_vl_n6_timeseries_2022.png"))
```

## n = 1

### Training data

```{r}
set.seed(13)
aug_vl_n1 <- augment(fit_vl_n1, data_n1_train)

aug_vl_n1 %>% ggplot(aes(log10_cases, .pred)) +
  geom_point() +
  stat_smooth(method = "lm") 

aug_vl_n1 %>% ggplot(aes(x=week)) +
  geom_line(aes(y=log10_cases), color = "blue") +
  geom_line(aes(y=.pred), color = "red")
```

### Testing data

```{r}
set.seed(13)
aug_vl_n1_test <- augment(fit_vl_n1, data_n1_test)

aug_vl_n1_test %>% ggplot(aes(log10_cases, .pred)) +
  geom_point() +
  stat_smooth(method = "lm")

aug_vl_n1_test %>% ggplot(aes(x=week)) +
  geom_line(aes(y=log10_cases), color = "blue") +
  geom_line(aes(y=.pred), color = "red")
```

# Comparing metrics

```{r}
cv_vl_n6_metrics <- cv_vl_n6_metrics %>% mutate(.config="vl_n6")
cv_vl_n5_metrics <- cv_vl_n5_metrics %>% mutate(.config="vl_n5")
cv_vl_n4_metrics <- cv_vl_n4_metrics %>% mutate(.config="vl_n4")
cv_vl_n3_metrics <- cv_vl_n3_metrics %>% mutate(.config="vl_n3")
cv_vl_n2_metrics <- cv_vl_n2_metrics %>% mutate(.config="vl_n2")
cv_vl_n1_metrics <- cv_vl_n1_metrics %>% mutate(.config="vl_n1")

cv_ap_n6_metrics <- cv_ap_n6_metrics %>% mutate(.config="df_n6")
cv_ap_n5_metrics <- cv_ap_n5_metrics %>% mutate(.config="df_n5")
cv_ap_n4_metrics <- cv_ap_n4_metrics %>% mutate(.config="df_n4")
cv_ap_n3_metrics <- cv_ap_n3_metrics %>% mutate(.config="df_n3")
cv_ap_n2_metrics <- cv_ap_n2_metrics %>% mutate(.config="df_n2")
cv_ap_n1_metrics <- cv_ap_n1_metrics %>% mutate(.config="df_n1")

cv_all_n6_metrics <- cv_all_n6_metrics %>% mutate(.config="all_n6")
cv_all_n5_metrics <- cv_all_n5_metrics %>% mutate(.config="all_n5")
cv_all_n4_metrics <- cv_all_n4_metrics %>% mutate(.config="all_n4")
cv_all_n3_metrics <- cv_all_n3_metrics %>% mutate(.config="all_n3")
cv_all_n2_metrics <- cv_all_n2_metrics %>% mutate(.config="all_n2")
cv_all_n1_metrics <- cv_all_n1_metrics %>% mutate(.config="all_n1")

cv_metrics <- rbind(cv_vl_n6_metrics,cv_vl_n5_metrics,cv_vl_n4_metrics,
                    cv_vl_n3_metrics,cv_vl_n2_metrics,cv_vl_n1_metrics,
                    cv_ap_n6_metrics,cv_ap_n5_metrics,cv_ap_n4_metrics,
                    cv_ap_n3_metrics,cv_ap_n2_metrics,cv_ap_n1_metrics,
                    cv_all_n6_metrics,cv_all_n5_metrics,cv_all_n4_metrics,
                    cv_all_n3_metrics,cv_all_n2_metrics,cv_all_n1_metrics)
```

## RSQ

```{r}
# Generate a list of 18 distinctive color hex codes
color_palette <- c(
  brewer.pal(8, "Dark2"), # 8 colors from the "Dark2" palette
  brewer.pal(9, "Set1"),  # 9 colors from the "Set1" palette
  "#66C2A5")              # Manually adding one more color for a total of 18

# Ensure the list has exactly 18 colors
color_palette <- color_palette[1:18]

# Make list for significance testing
comparisons <- list(c(1,6),c(2,6),c(6,12))

# Make list for naming models in ggplot
labels <- c("DF n=1","DF n=2","DF n=3","DF n=4","DF n=5","DF n=6",
            "VL n=1","VL n=2","VL n=3","VL n=4","VL n=5","VL n=6",
            "All n=1","All n=2","All n=3","All n=4","All n=5","All n=6")
```

```{r}
rsq_hist <- cv_metrics %>% filter(.metric=="rsq") %>% 
  ggplot(aes(.config, .estimate, fill=.config)) +
  geom_boxplot() +
  stat_compare_means(method = "anova", label.y = .9, label.x = 2) +
  stat_compare_means(aes(label = after_stat(p.signif)),
                  method = "t.test", ref.group = 6) +
  ggthemes::theme_clean() +
  theme(axis.text.x = element_blank(),
        axis.title = element_text(size = 8),
        legend.text = element_text(size = 6),
        legend.title = element_text(size = 8)) +
  scale_fill_manual(name="Configuration",
                    labels=labels,
                     values = color_palette) +
  xlab("Feature selection configuration") +
  ylab("RSQ value")

ggsave(here("figures/randforest_cv_rsq_boxplot.png"))
saveRDS(rsq_hist, here("figures/project_8040/rsq_histogram.rds"))

rsq_hist
```

```{r}
cv_select_ap_n6_metrics %>% ggplot(aes(.config,.estimate)) +
  geom_boxplot()
```

```{r}
cv_select_n6_metrics %>% ggplot(aes(.config,.estimate)) +
  geom_boxplot()
```

## RMSE

```{r}
rmse_hist <- cv_metrics %>% filter(.metric=="rmse") %>% 
  ggplot(aes(.config, .estimate, fill=.config)) +
  geom_boxplot() +
  ggthemes::theme_clean() +
  theme(axis.text.x = element_blank(),
        axis.title = element_text(size = 8),
        legend.text = element_text(size = 6),
        legend.title = element_text(size = 8)) +
  scale_fill_manual(name="Configuration",
                    labels=c("DF n=1","DF n=2","DF n=3","DF n=4",
                             "DF n=5","DF n=6","VL n=1","VL n=2",
                             "VL n=3","VL n=4","VL n=5","VL n=6",
                             "All n=6","All n=5","All n=4","All n=3",
                             "All n=2","All n=1"),
                     values = color_palette) +
  xlab("Feature selection configuration") +
  ylab("RMSE value") 

ggsave(here("figures/randforest_cv_rmse_boxplot.png"))
saveRDS(rmse_hist, here("figures/project_8040/rmse_histogram.rds"))

rmse_hist
```

# Summary stats

```{r}
data_n6_train %>% summarise_at(vars(14:28),mean)
data_n6_train %>% summarise_at(vars(14:28),median)
```

```{r}
cv_metrics_avg <- cv_metrics %>% group_by(.config,.metric) %>% 
  summarise(avg_estimate = mean(.estimate),
            sd_estimate = sd(.estimate))
```

```{r}
data_n6_train %>% ggplot(aes(log10_cases)) +
  geom_histogram(binwidth = .2)
```

```{r}
data_n6_train %>% ggplot(aes(cases.reported)) +
  geom_histogram(binwidth = 100)
```
